---
title: "ems_206_lab_5_vr"
author: "Vanessa Rathbone"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridata} to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heat with geom_tile()

# Attach packages

```{r, include = FALSE}

library(tidyverse)
library(here)
library(janitor)



```
```{r}

#Import Lobster data
# Case format = tidy format 
# Double check you don't have frequency table data (currently how lobster_count data is in))
# Time to tidy it with na = 

lobster_abundace <- read_csv(here::here("data", "lobster_abundance.csv"), 
                             na = "-99999") %>% 
  clean_names()



```

Use the tidyr::uncount() function to expand a frequency variable

```{r}

lobster_tidy <- lobster_abundace %>% 
  uncount(lobster_count)

```

Yey! Now each lobster has its own row. This is tidy format.

Let's look at this data in a bunch of different ways.

```{r}

ggplot(data = lobster_tidy, aes(x = site, 
                                y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2, 
              alpha = 0.3) #alpha is for transparacy

# Try a violin plot, good to look at datasets with large observations 

ggplot(data = lobster_tidy, aes(x = site, 
                                y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3) 
  

# Try histogram 

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free") # breaks out data into each group by argument, here it's by site

# Now a QQ plot

ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")

# In the qq plot, we observe AQUE is a relatively "linear" distribution. CARP and IVEE has a number of observations that are larger than expected for a totally normal sample distribution. MOHK and NAPL is relatively normally distributed too. Overall size_mm data is realtively normally distributed.  

```

Now, we're going to look at the data 
