---
title: "ems_206_lab_5_vr"
author: "Vanessa Rathbone"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridata} to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heat with geom_tile()

# Attach packages

```{r, include = FALSE}

library(tidyverse)
library(here)
library(janitor)



```
```{r}

#Import Lobster data
# Case format = tidy format 
# Double check you don't have frequency table data (currently how lobster_count data is in))
# Time to tidy it with na = 

lobster_abundace <- read_csv(here::here("data", "lobster_abundance.csv"), 
                             na = "-99999") %>% 
  clean_names()



```

Use the tidyr::uncount() function to expand a frequency variable

```{r}

lobster_tidy <- lobster_abundace %>% 
  uncount(lobster_count)

```

Yey! Now each lobster has its own row. This is tidy format.

Let's look at this data in a bunch of different ways.

```{r}

ggplot(data = lobster_tidy, aes(x = site, 
                                y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2, 
              alpha = 0.3) #alpha is for transparacy

# Try a violin plot, good to look at datasets with large observations 

ggplot(data = lobster_tidy, aes(x = site, 
                                y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3) 
  

# Try histogram 

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free") # breaks out data into each group by argument, here it's by site

# Now a QQ plot

ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")

# In the qq plot, we observe AQUE is a relatively "linear" distribution. CARP and IVEE has a number of observations that are larger than expected for a totally normal sample distribution. MOHK and NAPL is relatively normally distributed too. Overall size_mm data is realtively normally distributed.  

```

Use 'lubridate' to parse dates and times. 
Going to use mutate to add new column that has a date that is an actual date note a "character" classification
```{r}

# Make sure to specify the package when using 'lubridate'
# Can use class(lobster_tidy$date) in console to check how the date column is classified
# When changing the MM/DD/YYYY make sure to know the order of the date column, ours is in MDY
# Use lubridate to convert dates to "ISO 1806" date format YYYY-MM-DD

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date) #give column name in ()
  )

```

Parse the 'Date' column to get the different pieces (month, year) separated
```{r}

# By using lubridate::month function, it changes the class to "ordered" and "factor" so that all of the months are in order and known as a month factor instead of month as a charater (like if you use abb.numeric function)

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new))
```

Count lobsters by different groupings...

Let's say we want to ount the numbder of lobsters by year and month.
```{r}

# Can use group_by and summarize function, BUT can also use c'ount()'
# COUNT function says group_by, then summarize, then fine the length of something, and can also ungroup data (great for catagorical tidy data)

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month) #give it the variables that you want to group by, month and year here


# Just observations per year per site
lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)


# Just observations per site
lobster_site <- lobster_parse_date %>% 
  count(site)



# Now create summary statistics tables with multiple types of stats (mean, standard deviation, sample distribution)
# group_by and summarize and n()

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE), #use to remove any na values
    sd_size = sd(size_mm, na.rm = TRUE), #get's standard deviation
    sample_n = n()
  )


```


Confidence intervals and t-tests.

Use the 't-test()' function to find confidence intervals and perform t-tests
```{r}

# t distribution = use when you don't know the pop varience, accounts for uncertainty when dealing with smaller sample sizes (usually 30 or less)
# find Confidence intervals of observations at the IVEE site

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

# t-test to find confidence interval, read in console: mean lob size is significantly less than zero (p-value < 2.2e-16). Also shows only one sample is observed, number of lobsters
# single sample mean: 'mean of x 73.57968'  
# confidence interval is NOT about true pop range, IS about sample distribution means instead

t.test(ivee_lobsters) 

# OR if you don't wnat to filter and pull 

t.test(lobster_tidy$size_mm)

```

Two-sample t-test: 
Is there a significant difference in mean lobster lengths between Naples Reef and Mohawk Reef? 

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)

mohk_napl_ttest

# P-value (p-value < 2.2e-16) = as close to zero as the t-test will let you get
# there is only a 2.2e-16 probability that you could have taken 2 samples from populations and reached these 2 different 
# Infer that these sampels were drawn from populations that have different overall means, rather than the chance that we drew them from populations with similar means and by 2.2e-16 probability they were this different. (78.18454  66.86833)


```

INLINE REFERENCING 
Mean lobster size differed significantly between Mohawk and Naples reefs:
(t(`r mohk_napl_ttest$parameter`) = 19.849)...

Maybe an easier way to run a t.test with multiple observations
```{r}

lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

t.test_2 <- t.test(size_mm ~ site, data = lobster_2sample)

t.test_2
```

```{r}

# Heat map

ggplot(data = lobster_ysite, aes(x = obs_year, 
                                 y = site)) +
  geom_tile(aes(fill = n))

```

